\textbf{Lodging} \\
Suppose we have $n \geq 1$ rooms $r[1 \ldots n]$, and $m \geq 1$ groups $g[1 \ldots m]$. Each room $r[i]$ has a distinct ID
$r[i].id$ and a capacity $r[i].cap$, which is the maximum number of people that the room accommodates.
Each group g[i] has a distinct ID $g[i].id$ and a size $g[i].size$, which is the number of people in that group.
We want to assign groups to rooms so that each group is assigned to at most one room, at most one
group is assigned to each room, and if a group $g[i]$ is assigned to room $r[j]$, then $g[i].size \leq r[j].cap$.
Thus, an assignment A is a set of pairs $(g, r)$, where g is a group, $r$ is a room, and the following conditions
are satisfied: (i) If $(g, r)$ and $(g', r)$ are distinct pairs in A, then $g.id  \ne g'.id$ and $r.id  \ne r'.id$; and (ii)
If $(g, r) \in A$, then $g.size \leq r.cap$.\\

Notice that if there are fewer rooms than groups or if rooms are not big enough, not all groups can be
assigned rooms. Similarly, if there are too many rooms, some rooms might go empty (i.e., no groups
are assigned to these rooms). \\

The merit of an assignment is the sum of the sizes of the groups that are assigned rooms. Give a
greedy algorithm that prints an assignment of maximum merit. Be sure to rigorously state and prove
the Greedy Choice Lemma that your algorithm is based on. State the time and space complexity of
your algorithm

\begin{customsolutionbox}
    \underline{\textbf{Greedy choice lemma}}: Given a set of rooms $R = \{r_1, r_2, \ldots, r_n\}$ and a set of groups $G = \{g_1, g_2, \ldots, g_m\}$, let 
    $g^* \in G$ be the largest unassigned group (i.e., with the maximum size), and let $r^* \in R$ be the largest available room such that 
    $g^*.size \leq r^*.cap$. Then there exists an optimal assignment in which the pair $(g^*, r^*)$ is included, and for some room $r' \leq r^*$, we can find 
    an optimal assignment $(g', r')$ where $g' \leq g^*$. 

    \begin{proof}
        Let $S$ be the largest set with the maximum merit. Consider the following cases:
        \begin{parts}
            \part If $(g^*, r^*) \notin S$, then we can define $S' = S \cup \{(g^*, r^*)\}$, yielding $|S'| > |S|$. This contradicts that $S$ is the largest set with the maximum possible merit,
            and $S'$ os the largest possible set with maximum merit.
    
            \part Suppose $(g', r^*) \in S$ and $(g^*, r^*) \notin S$. Since $g^*.size \leq r^*.cap$ by assumption, we can define 
            \[
            S' = \left(S \setminus \{(g', r^*)\}\right) \cup \{(g^*, r^*)\}
            \]
            where $|S'| = |S|$. Therefore, $S'$ is also a largest set with maximum merit.
    
            \part Suppose $(g^*, r') \in S$ and $(g^*, r^*) \notin S$. Since $r^* \leq r'$ in capacity, we can define 
            \[
            S' = \left(S \setminus \{(g^*, r')\}\right) \cup \{(g^*, r^*)\}
            \]
            where $|S'| = |S|$. Therefore, $S'$ is also a largest set with maximum merit.
    
            \part Suppose we have the pairs $(g', r^*)$ and $(g^*, r')$ in $S$. Then we can create the pairs $(g^*, r^*)$ and $(g', r')$, and define
            \[
            S' = \left(S \setminus \{(g', r^*), (g^*, r')\}\right) \cup \{(g^*, r^*), (g', r')\}
            \]
            so that $|S'| = |S|$. Thus, $S'$ is also a largest set with maximum merit.
        \end{parts}
    \end{proof}
    

    \begin{algorithm}[H]
        \caption{LodgingProblem}
        \SetKwFunction{FMaximumMerit}{MaximumMerit}
        \SetKwProg{Fn}{Function}{:}{}
        \Fn{\FMaximumMerit($r[1 \ldots n], g[1 \ldots m]$)}{
            Sort $r[1 \ldots n]$ descending by $r[i].cap$ \\
            Sort $g[1 \ldots m]$ descending by $g[i].size$ \\
            $ptr \gets 1$ \tcp*{Pointer to next unassigned group}
            maxMerit $\gets 0$ \\
    
            \For{$i \gets 1$ \KwTo $n$}{
                \While{$ptr \leq m$ \textbf{and} ($g[ptr].size >  r[i].cap$)}{
                    ptr++
                }   
                \If{$ptr \leq m$}{
                    maxMerit $\gets$ maxMerit g[ptr].size \\
                    ptr++
                }   
            }
    
            \KwRet maxMerit
        }
    \end{algorithm}
    

    \underline{Time complexity}: $O(n log n + m log m)$ \\
    \underline{Space complexity}: $O(n+m)$
\end{customsolutionbox}