\textbf{A Reflection on Algorithms} \\
\begin{questions}
    \question Importance of the foundations of algorithms and theory of computation \\
       I believe the foundations of algorithms are essential for understanding efficient problem-solving techniques, and they are especially necessary 
       when building complex systems that process large data sets or serve many users concurrently. While I don't have much experience with the theory of computation—largely due 
       to the pivots I've made in my educational path at Dartmouth—I recognize its value. It provides the theoretical underpinnings for what can be computed and how efficiently. 
       This is crucial for designing algorithms that are not only correct but also optimal in terms of time and space complexity. Overall, I believe algorithms 
       help automate redundant tasks and enable us to focus on more complex and creative aspects of problem-solving.
    
    \question The importance of algorithms in your everyday life. \\
        Algorithms play a significant role in my everyday life, even if I don't always recognize them. From the way my smartphone organizes photos to how social media platforms 
        recommend content, algorithms operate behind the scenes. They help me find information quickly, automate routine tasks, and even optimize daily activities. 

        In my academic work, algorithms help me analyze complex biological data—such as tracking pixel intensity changes in fluorescence microscopy images—to understand how cells take up 
        imaging agents. This process involves a combination of physics-based models, biological computations, and image processing algorithms. Together, they enable deeper insights that 
        would be difficult to achieve manually.

    \question The (expected) importance of algorithms in your career \\
        In the near future, I will be working as a software engineer, and a strong understanding of algorithms—especially legacy and foundational ones—will be essential. 
        Algorithms will be fundamental to writing efficient code, debugging existing systems, and designing scalable solutions. Whether optimizing some infrastructure or handling 
        large-scale data, algorithmic thinking will be central to my career.

    \question The (expected) importance of algorithms in your career \\
        Personally, I find mathematical reasoning challenging at first, but I've come to appreciate how algorithms bring structure to problem-solving. 
        As I prepare for a career in software engineering, I expect algorithms to be central to my work—whether it's improving performance, ensuring scalability, or 
        understanding how systems behave under the hood. Even when the math is tough, it's rewarding to see how those principles power real-world applications.
 
\end{questions}
