\textbf{Printing} \\
Consider the problem of neatly printing a paragraph with a monospaced font (all characters having the
same width) on a printer. The input text is a sequence of n words of lengths $\ell_1, \ell_2, . . . , \ell_n$, measured
in characters. We want to print this paragraph neatly on a number of lines that hold a maximum of
$M$ characters each. Our criterion of “neatness” is as follows. If a given line contains words $i$ through
$j$, where $i \leq j$ , and we leave exactly one space between words, the number of extra space characters
$\ell_k$ , which must be nonnegative so that the words fit on the
line. We wish to minimize the sum, over all lines except the last, of the cubes of the numbers of extra
Design a dynamic programming algorithm of $O(n^2)$ time complexity and $O(n)$ space complexity that,
given $n \geq 1$, $\ell[1 . . . n]$, and $M \geq 1$, prints neatly a paragraph of n words of lengths given by the $\ell$ array.

\begin{customsolutionbox}
    \underline{Clever observation}: 
        For words up to index $i$, we consider all possible break points at index $j$ ($1 \leq j \leq i$), where words $j$ through $i$ would be placed on the current line. 
        For each possible $j$, we compute the penalty of placing words $j$ to $i$ on one line (or $\infty$ if they do not fit). 
        The dynamic programming step chooses the $j$ that minimizes the total cost up to $i$, which consists of the optimal cost up to $j-1$ plus the penalty for words $j$ to $i$. 
        The corresponding $j$ is stored in $breakPoint(i)$ to allow reconstruction of the optimal paragraph.


    \underline{Notation}: We we have the following:
    \begin{itemize}
        \item extraSpaces($j, i$) = $M - i+j - \sum_{r=j}^k \ell_r$
        \item breakPoint($j$) - position with minimum penalty for printing words through some indices $j$ to $i$ for $1 \leq j \leq i \leq n$ 
        \item penalty($j, i$)
            \[
                \text{penalty}(j, i) =
                \begin{cases}
                0, & \text{if } i = n \ (\text{last line has no penalty}) \\[8pt]
                \left(\text{extraSpaces}(j, i)\right)^3, & \text{if } \text{extraSpaces}(j, i) \geq 0 \\[8pt]
                \infty, & \text{if } \text{extraSpaces}(j, i) < 0
                \end{cases}
            \]
        \item costs($j, i$) - minimizes the costs of printing a sequence of characters in a new line.
    \end{itemize}

    \begin{algorithm}[H]
        \caption{PrintNeatly}
        \underline{Preconditions}: $n \geq 1, M \leq 1$, $\ell[1 \ldots n]$ is array of real numbers representing lengths of words. \\
        \underline{Postcondition}: Terminates, and prints the paragraph of $n$ words neatly \\
        \SetKwFunction{FPrintNeatly}{PrintNeatly}
        \SetKwProg{Fn}{Function}{:}{}
        \Fn{\FPrintNeatly($\ell[1 \ldots n], M$)}{
            totalChars$[1] \gets \ell_1$ \\
            \For{$i \gets 2$ \KwTo $n$} {
                totalChars$[i] \gets$ totalChars[$i-1$] + $\ell[i]$
            }
            \vspace{1em}
            penalty $\gets 0$ \\
            breakPoints $\gets []$ \\
            Initialize all costs[$1 \ldots n]$ with $\infty$ \\
            \For{$i \gets 1$ \KwTo $n$}{
                \For{$j \gets i$ \KwTo $1$}{
                    charLength $\gets$ totalChars$[i]$ - totalChars[$j-1]$ \\
                    spaces $\gets i - j$ \\
                    extraSpaces $\gets M$ - charLength - spaces \\

                    \If{extraSpaces $< 0$}{
                        penalty $\gets \infty$
                    }

                    \If{$i = n$}{
                        penalty $\gets 0$
                    }
                    \Else{
                        penalty $\gets$ (extraSpaces)$^3$
                    }

                    \If{costs$[j-1] +$ penalty $<$ costs$[i]$}{
                        costs$[i] \gets$ costs$[j-1] +$ penalty \\
                        breakPoints$[i] \gets j$
                    }
                }
            }
            printParagraph($1, n$, breakPoints)
        }
        \vspace{2em}
        \underline{Preconditions}: $\ell, n \in \N, \ell \leq n$, breakPoints$[\ell \ldots n]$ is array of real numbers, where $\ell$ is the print starting point \\
        \underline{Postcondition}: Terminates, and prints the paragraph neatly \\
        \SetKwFunction{FPrintParagraph}{PrintParagraph}
        \Fn{\FPrintParagraph(start, $n$, breakPoints)}{
            \If{ start $> n$}{ \KwRet }

            endIndex $\gets$ breakPoints$[start]$ \\
            prints words $start$ through $endIndex$ on line, then newline \\
            \FPrintParagraph($endIndex+ 1, n, breakPoints$)
        }
    \end{algorithm}

    \vspace{0.5em}
    \underline{Time complexity}: We have $O(n^2) + O(n)$, hence the overall time complexity is $O(n^2)$ \\
    \underline{Space complexity}: We have $O(3n)$ because of the 3 arrays to used hence $O(n)$ \\
\end{customsolutionbox}