\text{Semiconnectivity} \\
A directed graph $G= (V,E)$ is semiconnected if, for all pairs of vertices $u,v \in V$, there is a path from
$u$ to $v$ or there is a path from $v$ to $u$(or both). Give an efficient algorithm that, given a directed graph
$G= (V,E)$, outputs whether or not it is semiconnected. Argue the correctness of your algorithm and
analyze its time complexity to be $\Theta(V + E)$

\begin{customsolutionbox}
    \begin{algorithm}[H]
        \caption{Semiconnectivity Check}
        \SetKwFunction{FSemiconnected}{Semiconnected}
        \SetKwFunction{FKosaraju}{Kosaraju}
        \SetKwFunction{FComponentGraph}{ComponentGraph}
        \SetKwProg{Fn}{Function}{:}{}
        \Fn{\FSemiconnected{$G(V, E)$}}{
            $(k, S) \gets$ FKosaraju{$G$} \textbf{ // SCC ID array; from q1} \\
            $G^{SCC} \gets$ FComponentGraph{$G, S$} \textbf{ // from q3} \\

            Do topological sort on $G^{SCC}$ \\

            \For{$i \gets 1$ \KwTo $k-1$} {
                $C_i \gets$ the $i^{th}$ SCC in the topological order \\
                $C_{i+1} \gets$ the $(i+1)^{th}$ SCC in the topological order \\
                \If{$C_{i+1} \not\in A^{SCC}[C_i]$} {
                    \Return \textbf{false}
                }
            }

            \Return \textbf{true}
        }
    \end{algorithm}

    \textbf{Correctness}: The algorithm first identifies SCCs because semiconnectivity is preserved within SCCs. To check for semiconnectivity of 
    the SCCs, we construct the component graph $G^{SCC}$ and perform a topological sort. We then verify that each SCC can reach the next one in the 
    topological order. If any SCC cannot reach the next, the graph is not semiconnected. \\

    \textbf{Time Complexity}: The algorithm runs in $\Theta(V + E)$ time. The Kosaraju's algorithm for finding SCCs runs in $\Theta(V + E)$, and 
    constructing the component graph also takes $\Theta(V + E)$. The topological sort of the component graph takes $\Theta(V + E)$ as well. Thus, 
    the overall time complexity is $\Theta(V + E)$. \\
    
    \textbf{Space Complexity}: The space complexity is also $\Theta(V + E)$ due to the storage of the adjacency list for the component graph and the 
    SCC ID array.

\end{customsolutionbox}
